---
output: rmarkdown::github_document
editor_options: 
  chunk_output_type: console
---
```{r pkg-knitr-opts, include=FALSE}
hrbrpkghelpr::global_opts()
```

```{r badges, results='asis', echo=FALSE, cache=FALSE}
hrbrpkghelpr::stinking_badges()
```

# ulid

Universally Unique Lexicographically Sortable Identifiers

## Description

(grifted from <https://github.com/ulid/spec>)

UUID can be suboptimal for many uses-cases because:

- It isn't the most character efficient way of encoding 128 bits of randomness
- UUID v1/v2 is impractical in many environments, as it requires access to a unique, stable MAC address
- UUID v3/v5 requires a unique seed and produces randomly distributed IDs, which can cause fragmentation in many data structures
- UUID v4 provides no other information than randomness which can cause fragmentation in many data structures

Instead, herein is proposed ULID:

```javascript
ulid() // 01ARZ3NDEKTSV4RRFFQ69G5FAV
```

- 128-bit compatibility with UUID
- 1.21e+24 unique ULIDs per millisecond
- Lexicographically sortable!
- Canonically encoded as a 26 character string, as opposed to the 36 character UUID
- Uses Crockford's base32 for better efficiency and readability (5 bits per character)
- Case insensitive
- No special characters (URL safe)
- Monotonic sort order (correctly detects and handles the same millisecond)

```
 01AN4Z07BY      79KA1307SR9X4MV3

|----------|    |----------------|
 Timestamp          Randomness
   48bits             80bits
```

### Components

**Timestamp**
- 48 bit integer
- UNIX-time in milliseconds
- Won't run out of space till the year 10889 AD.

**Randomness**
- 80 bits
- Cryptographically secure source of randomness, if possible

### Sorting

The left-most character must be sorted first, and the right-most character sorted last (lexical order). The default ASCII character set must be used. Within the same millisecond, sort order is not guaranteed.

## What's Inside The Tin

The following functions are implemented:

```{r ingredients, results='asis', echo=FALSE, cache=FALSE}
hrbrpkghelpr::describe_ingredients()
```

## Installation

```{r install-ex, results='asis', echo=FALSE, cache=FALSE}
hrbrpkghelpr::install_block()
```

## Usage

```{r message=FALSE, warning=FALSE, error=FALSE}
library(ulid)

# current verison
packageVersion("ulid")

```

### One

```{r}
ulid::ULIDgenerate()
```

### Many

```{r}
(u <- ulid::ULIDgenerate(20))
```

### Unmarshal

```{r}
unmarshal(u)
```

### Use defined timestamps

```{r}
(ut <- ts_generate(as.POSIXct("2017-11-01 15:00:00", origin="1970-01-01")))

unmarshal(ut)
```

## ulid Code Metrics

```{r}
cloc::cloc_pkg_md()
```

## Code of Conduct

Please note that this project is released with a Contributor Code of Conduct. By participating in this project you agree to abide by its terms.
